### 2025-07-19 evaluating-prediction-accuracy.qmd
20250719 - ROC curve

When dealing with a categorical outcome, discrimination is the ability to separate events from non-events.
When dealing with a continuous outcome, discrimination is the strength of the association between the predictor and the outcome.
Aspects of discrimination at a particular cutoff (e.g., sensitivity, specificity, area under the ROC curve) are described in INSERT.

@@

@@

INSERT depicts an empirical ROC plot where lines are drawn to connect the hit and false alarm rates.

INSERT depicts an ROC curve where a smoothed and fitted curve is drawn to connect the hit and false alarm rates.

@@

To generate a 2x2 confusion matrix, we need to make a dichotomous decision/prediction.
For this example, our goal is to predict which Running Backs or Wide Receivers will score 300 or more fantasy points.
We will predict a player as scoring 300 or more fantasy points for a given projection source if that projection source projected the player to score 300 or more fantasy points.
Thus, the same player has multiple rows in the data file—one for each projection source.

>>>>

When dealing with a categorical outcome, discrimination is the ability to separate events from non-events.
When dealing with a continuous outcome, discrimination is the strength of the association between the predictor and the outcome.
Aspects of discrimination at a particular cutoff (e.g., sensitivity, specificity, area under the ROC curve) are described in @sec-thresholdDependentAccuracy.

@@

### Set a Cutoff/Threshold {#sec-setCutoff}

For evaluating sensitivity, specificity, positive predictive value, negative predictive value, and the [receiver operating characteristic (ROC) curve](#sec-roc), we need to make a dichotomous decision/prediction.
For this example, our goal is to predict which Running Backs or Wide Receivers will score 300 or more fantasy points.
We will predict a player as scoring 300 or more fantasy points for a given projection source if that projection source projected the player to score 300 or more fantasy points.
Thus, our cutoff or threshold is 300 or more projected points.
The same player has multiple rows in the data file—one for each projection source.

@@

@fig-aucExampleEmpirical depicts an empirical ROC plot where lines are drawn to connect the hit and false alarm rates, in predicting Running Backs and Wide Receivers who score 300 or more fantasy points.

```{r}
#| label: fig-aucExampleEmpirical
#| fig-cap: "Empirical Receiver Operating Characteristic (ROC) Curve for Predicting Running Backs and Wide Receivers that Score 300 or More Fantasy Points."
#| fig-alt: "Empirical Receiver Operating Characteristic (ROC) Curve for Predicting Running Backs and Wide Receivers that Score 300 or More Fantasy Points."
#| fig-width: 8
#| fig-height: 8

plot(
  pROC::roc(
    truth ~ raw_points,
    projectionsWithActuals_seasonal_subset),
  legacy.axes = TRUE,
  print.auc = TRUE)
```

@fig-aucExample depicts an ROC curve where a smoothed and fitted curve is drawn to connect the hit and false alarm rates, in predicting Running Backs and Wide Receivers who score 300 or more fantasy points.

```{r}
#| label: fig-aucExample
#| fig-cap: "Smoothed Receiver Operating Characteristic (ROC) Curve for Predicting Running Backs and Wide Receivers that Score 300 or More Fantasy Points."
#| fig-alt: "Smoothed Receiver Operating Characteristic (ROC) Curve for Predicting Running Backs and Wide Receivers that Score 300 or More Fantasy Points."
#| fig-width: 8
#| fig-height: 8

plot(
  pROC::roc(
    truth ~ raw_points,
    projectionsWithActuals_seasonal_subset,
    smooth = TRUE),
  legacy.axes = TRUE,
  print.auc = TRUE)
```

@@

Here is our 2x2 confusion matrix, predicting which Running Backs or Wide Receivers will score 300 or more fantasy points.
